---
title: "Working_with_Data"
format: 
  pdf:
    number-sections: true
    toc: true
    pdf-engine: xelatex 
    geometry: margin=1in
    fontsize: 8pt
editor: visual
---

## Packages & library

```{r library}
library(tidyverse)
library(descr)
library(knitr)
library(dplyr)
library(Hmisc)
library(readr)
library(readxl)
library(ggplot2)
```

## Load

```{r load-data}
Project_Data = read.csv("/cloud/project/Data/Connection_to_Nature_Data.csv",
header = TRUE)
```

## Variables

```{r}
# VARIABLE 1: People's Age


# Age 18 and above

Project_Data <- subset(Project_Data, D_Age >= 18)

# Group D_Age into 4 age range's and Labeling

Project_Data$Age_Group <- cut(Project_Data$D_Age, 
                              breaks = c(18, 25, 40, 65, Inf), 
                              labels = c("18-25", "26-40", "40-65", "65+"),
                              right = TRUE,
                              include.lowest = TRUE)

# Frequency table 
                              
freq(as.ordered(Project_Data$Age_Group),plot =FALSE)
```

I choose this variable (age) because I think it would be important to look at in reference to how loneliness and time spent in nature varies among age groups. Perhaps depending on the age group, there will be more positive benefits to those exposed to nature in relation to loneliness.

```{r}
# VARIABLE 2: How many hours on average do you currently spend in nature per week? 

# Group D_hours into 4 categories & Labeling

Project_Data$Nature_Hours_Group <- cut(Project_Data$D_hours, 
                                       breaks = c(0, 5, 15, 30, Inf),
                                       labels = c("Low (0-5)", "Moderate (6-15)", "High (16-30)", "Very High (30+)"), 
                                       include.lowest = TRUE)

# Frequency table

freq(as.ordered(Project_Data$Nature_Hours_Group), plot = FALSE)
```

This is very important. This variable (hours spent in nature) is important because when I did the literature review assignment, depending on the time spent in nature, actually lowered both social loneliness and emotional loneliness, but it depends how much time was spent in nature.

```{r}
# VARIABLE 3: People's experience a general sense of emptiness (survey response)

# Labeling

Project_Data$Lon_1 <- factor(Project_Data$Lon_1,
                             levels = c(1, 2, 3),
                             labels = c("yes", "more or less", "no"))

# Frequency table

freq(as.ordered(Project_Data$Lon_1), plot = FALSE)

# Bar graph

ggplot(data = subset(Project_Data, !is.na(Lon_1)), aes(x = Lon_1)) +
  geom_bar(fill = "red", color = "black") +
  xlab("Self-Reported Emptiness") +
  ylab("Survey Count") +
  ggtitle("Do People Experience a general sense of Emptiness?") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, face = "bold"))
```

-   Lit Review Assign: I choose this variable (people's sense of emptiness) because this can be a reason for social/emotional loneliness. If time spent nature is associated with lower loneliness on these two paths, we might also see a decrease in emptiness to those who spend more time in nature.
-   Univariate Data Visualization Assign: The first graph illustrates people's experiences of a general sense of emptiness. Overall, most respondents said no, while the fewest number of respondents said yes. 

```{r}
# VARIABLE 4: I miss having people around (survey response)

# Labeling

Project_Data$Lon_4 <- factor(Project_Data$Lon_4,
                             levels = c(1, 2, 3),
                             labels = c("yes", "more or less", "no"))

# Frequency table

freq(as.ordered(Project_Data$Lon_4), plot = FALSE)

# Bar graph

ggplot(data = subset(Project_Data, !is.na(Lon_4)), aes(x = Lon_4)) +
  geom_bar(fill = "red", color = "black") +
  xlab("Feeling About Missing People") +
  ylab("Survey Count") +
  ggtitle("Do People Feel They Miss Having Others Around?") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, face = "bold"))
```

-   Lit Review Assign: This variable (missing social interaction) could be important because social loneliness is being examined here. Comparing this to time spent in nature can help show whether nature can also regulate/help social loneliness as well.

-   Univariate Data Visualization Assign: The second graph illustrates people's survey responses to whether individuals feel that they miss others in their lives. This is similar to the first graph. Most respondents said no, while the fewest number of respondents said yes. 

```{r}
# VARIABLE 5: I have high self-esteem (survey response)

# Labeling

Project_Data$SE_1 <- factor(Project_Data$SE_1,
                            levels = c(1, 2, 3, 4, 5),
                            labels = c("not very true of me", "2", "3", "4", "very true of me"),
                            ordered = TRUE)

# Frequency table 

freq(as.ordered(Project_Data$SE_1), plot = FALSE)
```

Lastly, I also choose this variable (people's self-esteem) because those who experience loneliness and spend little time in nature differ from those who don't feel loneliness and do spend time in nature. Perhaps those who do spend more time have higher level's of agreement to self-esteem compared to those who do not.

## Bivariate Analysis

```{r}
# Bivariate 1: Correlation between age and time spent in nature

Project_Data <- subset(Project_Data, !is.na(D_Age) & !is.na(D_hours))

# Scatter plot 

ggplot(Project_Data, aes(x = D_Age, y = D_hours)) +
  geom_point(color = "orange", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  xlab("Age (Years)") +
  ylab("Hours Outside per Week") +
  ggtitle("Does Age Influence Time Spent in Nature?") +
  theme_classic() + 
  theme(plot.title = element_text(size = 14, face = "bold"))  
```

-   The scatter plot shown here shows that there is a slight positive correlation between age and time spent outside hourly per week. The positive correlation shows that older individuals might spend a little bit more time in nature. However, when it comes to my research interest concerning the relationship between loneliness and nature, age does not seem to be a strong factor.

```{r}
# Bivariate 2: Correlation between time spent in nature and general emptiness  

summary_data <- Project_Data %>%
  filter(!is.na(Lon_1) & !is.na(D_hours)) %>%
  group_by(Lon_1) %>%
  summarise(Avg_Hours = mean(D_hours))

# Bar graph
ggplot(summary_data, aes(x = Lon_1, y = Avg_Hours, fill = Lon_1)) +
  geom_bar(stat = "identity", fill = "red", color = "black") +
  xlab("Sense of Emptiness") +
  ylab("Average Hours Outside per Week") +
  ggtitle("Average Hours in Nature by Sense of Emptiness") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, face = "bold")) +
  guides(fill = FALSE)
```

-   The bar graph shown here illustrates the relationship between people's sense of emptiness, and their average hours spent in nature. Individuals who say YES to feeling empty, show, a high number of hours outside, and are like those who say NO to feeling empty as well. However, people who said MORE OR LESS, spent the least amount of time outside. They would mean that the relationship between a sense of emptiness and time spent in nature is not straightforward and there could be other factors influencing nature and loneliness.
