---
title: "ANOVA One Way"
format: 
  pdf:
    number-sections: true
    toc: true
    pdf-engine: xelatex 
    geometry: margin=1in
    fontsize: 8pt
editor: visual
---

```{r}
data("ToothGrowth")
```

```{r}
library(car)
```

```{r}
# Convert dose to a factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# ANOVA One-Way
anova_result <- aov(len ~ dose, data = ToothGrowth)
summary(anova_result)

# Assumption check
leveneTest(len ~ dose, data = ToothGrowth)

# Box-plot
library(ggplot2)
ggplot(ToothGrowth, aes(x = dose, y = len)) +
  geom_boxplot(fill = "violet") +
  labs(title = "Tooth Length by Vitamin C Dose",
    x = "Dose (mg/day)", y = "Tooth Length") +
    theme_classic()
```

```{r}
#Post-Hoc Test
TukeyHSD(anova_result)
```

Null Hypothesis (H0): There is no difference in the averages among the three different Vitamin C dose groups (0.5 mg, 1 mg, and 2 mg).

Alternative Hypothesis (H1): There is a difference (at least two) in the average tooth length among the dose groups.

In the ANOVA one-way test, I checked to see if different amounts of Vitamin C affected how much tooth grew for guinea pigs. The test showed a statistically significant difference between the groups, meaning that at the very least (potentially), one of the dose levels led to a different amount of tooth growth.

The boxplot also shows a visible difference in the median tooth length between the three dose groups.

Therefore, I reject the null hypothesis. Vitamin C does affect guinea pig's tooth growth.

At the post-hoc test, I found that all three comparisons were statistically significant, with higher doses having greater tooth growth.

\*\*Only Vitamin C was looked upon, but not Orange Juice from the dataset.
